---
title: "Assignment 2"
author: "Mason Rocco"
date: "2023-02-09"
output: pdf_document
editor_options: 
markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 2

### Question 1

#### Part a

\
```{r warning=FALSE, message=FALSE}
library(readr)
library(tidyverse)
hiroshima <- read_table("C:/Users/mrocc/Desktop/STAT 3302/Data/hiroshima.txt")
hiroshima$prop <- hiroshima$leukemia / hiroshima$total

ggplot(hiroshima, aes(x = midpoint, y  = prop)) + 
  labs(x= "Radiation Level (Radons)", y = "Proportion of Deaths due to Leukemia",
  title = "Proportion of Cancer Deaths Due to Leukemia vs. Radiation Level") +
  geom_point()
```


Looking at the graph, we can see that the proportion of cancer deaths due to Leukemia increases as the radiation level increases.
\

#### Part b

\

```{r warning=FALSE, message=FALSE}
death_1 <- c(rep(1, 13), rep(0, 378))
mid_1 <- rep(0, 391)
death_2 <- c(rep(1, 5), rep(0, 200))
mid_2 <- rep(5, 205)
death_3 <- c(rep(1, 5), rep(0, 151))
mid_3 <- rep(29.5, 156)
death_4 <- c(rep(1, 3), rep(0, 47))
mid_4 <- rep(74.5, 50)
death_5 <- c(rep(1, 4), rep(0, 31))
mid_5 <- rep(149.5, 35)
death_6 <- c(rep(1, 18), rep(0, 33))
mid_6 <- rep(249.5, 51)
deaths <- c(death_1, death_2, death_3, death_4, death_5, death_6)
midpoints <- c(mid_1, mid_2, mid_3, mid_4, mid_5, mid_6)

hiroshima_Bern <- glm(formula = deaths ~ midpoints, 
                       data = data.frame(deaths, midpoints), family = "binomial")
summary(hiroshima_Bern)
```

```{r warning=FALSE, message=FALSE}
hiroshima_Bin <- glm(prop~midpoint, data=hiroshima, family="binomial", weight=total)
summary(hiroshima_Bin)
```
\
Using these two models, we can see that a one unit increase in radiation results in the logit of the probability of a cancer death being leukemia by approximately 0.012. 

$$
\frac{odds(x_i+1)}{odds(x_i)} = 
\frac{exp[\beta_0+\beta_1x_i+\beta_1]}{exp[\beta_0+\beta_1x_i]} = e^{\beta_1}
$$

For a one unit increase in radiation, we see that the odds of a cancer death being leukemia increases approximately 1.01 times, or by 1.117%.

A 95% Confidence interval for $\hat\beta_1$ is as follows: 
$0.011624 \pm 1.96 * 0.001487$ = (0.0087, 0.0145)

### Question 2

#### Part a
\
$p_i (i=1,...,n)$ 
\
$x_i (i=1,...,n)$
\
$log(\frac{p_i}{1-p_i}) = \beta_0+\beta_1x_i$
\
$p_i = \frac{e^{\beta_0+\beta_1x_i}}{1+e^{\beta_0+\beta_1x_i}}$ 
\

Where $\beta_0$ is the probability that a patient has a sore throat after a surgery that is 0 minutes long, $\beta_1$ is the change in probability that a patient has a sore throat for each added minute of surgery, $p_i$ is the probability that the patient has a sore throat after surgery, and $x_i$ is the length of the surgery in minutes.

### Part b

```{r warning=FALSE, message=FALSE}
throat <- read_table("C:/Users/mrocc/Desktop/STAT 3302/Data/throat.txt")

throat_model <- glm(formula = throat ~ duration, data = throat, family = "binomial")
summary(throat_model)
```


This tells us that a one minute increase in the duration of surgery increases 
the logit of the probability that a patient will have a sore throat by 0.07203.

$$
\frac{odds(x_i+1)}{odds(x_i)} = 
\frac{exp[\beta_0+\beta_1x_i+\beta_1]}{exp[\beta_0+\beta_1x_i]} = e^{\beta_1}
$$

Thus, we can see that a one minute increase in the surgery duration results in an
estimated sore throat by 1.07028 times, or an increase of 7.028%.

A 95% Confidence interval for $\hat\beta_1$ is as follows: 
$0.07203 \pm 1.96 * 0.02545$ = (0.0221, 0.1219)

