---
title: "Assignment 3"
author: "Mason Rocco"
date: "2023-03-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library("nnet")
```


## Problem 1
### a.
Model: $\beta_{0j}+\beta_{1j}x, j=1,2$

Fitted Model:
\
Where x = income level
\
"Very Happy" = 3 (Reference), "Pretty Happy" = 2, "Not Happy" = 1
\
$log(\frac{\hat{p_1}}{\hat{p_3}}) = \hat\beta_{01} + \hat\beta_{11}x = -2.555 - 0.228x$
\
$log(\frac{\hat{p_2}}{\hat{p_3}}) = \hat\beta_{02} + \hat\beta_{12}x = -0.351 - 0.096x$

### b.
$e^{-0.2277717} = 0.796$ 
\
For each additional level of income, the "chance" that a person is "not happy" 
(vs. "very happy") in their marriage decreases 0.796 times.

### c.
$\hat\beta_{12} = -0.096$
\
$e^{-0.096} = 0.908$
\
$e^{\hat\beta_{12}}\pm1.96*SE(\hat\beta_{12}) = e^{-0.096}\pm1.96*0.122 = (0.715, 1.154)$
\
When creating a 95% confidence interval for the odds ratio of $\hat\beta_{12}$, 
we note that the interval includes 1, meaning that there is not a significant 
difference between "very happy" and "pretty happy," when adjusting for income level.
### d.
```{r}
# Income Level 2 model-predicted probability ("Very Happy")
p3 <- 1/(1+exp(-0.3512438-(2*-0.0961703))+exp(-2.5545786-(2*-0.2277717)))
p3
```

```{r}
# Income Level 2 model-predicted probability ("Pretty Happy")
p2 <- p3 * exp(-0.3512438-(2*-0.0961703))
p2
```

```{r}
# Income Level 2 model-predicted probability (Not Happy")
p1 <- p3 * exp(-2.5545786-(2*-0.2277717))
p1
```


## Problem 2
### a.
```{r}
col1 <- c(rep("Sequential", 151), rep("Alternating", 148))
col2 <- c(rep("Male", 128), rep("Female", 23), rep("Male", 125), rep("Female", 23))
col3 <- c(rep("Progressive Disease", 28), rep("No Change", 45), 
rep("Partial Remission", 29), rep("Complete Remission", 26), rep("Progressive Disease", 4),
rep("No Change", 12), rep("Partial Remission", 5), rep("Complete Remission", 2),
rep("Progressive Disease", 41), rep("No Change", 44), rep("Partial Remission", 20),
rep("Complete Remission", 20), rep("Progressive Disease", 12), rep("No Change", 7),
rep("Partial Remission", 3), rep("Complete Remission", 1))
data <- data.frame(col1, col2, col3)
table(data$col3, data$col2, data$col1)
```

### b.
```{r}
col1_2 <- c(rep(0, 151), rep(1, 148))
col2_2 <- c(rep(1, 128), rep(0, 23), rep(1, 125), rep(0, 23))
col3_2 <- c(rep(1, 28), rep(2, 45), rep(3, 29), rep(4, 26),
rep(1, 4), rep(2, 12), rep(3, 5), rep(4, 2),
rep(1, 41), rep(2, 44), rep(3, 20), rep(4, 20),
rep(1, 12), rep(2, 7), rep(3, 3), rep(4, 1))
freq.table2 <- data.frame(col1_2, col2_2, col3_2)
colnames(freq.table2) <- c("Alternating", "Male", "DiseaseChange")
freq.table2$response.fac <- factor(freq.table2$DiseaseChange, levels=c(1,2,3,4))
summary(polr(response.fac ~ Male + Alternating, data=freq.table2))
```

Treatment (Alternating) Factor Coefficient: -0.5807
Odds: $e^{-0.581} = 0.5595$
This means that a patient who received alternating treatment have odds of being 
in a higher category of disease progression that are 0.5595 times lower than 
that of a patient who received sequential treatment, adjusting for sex.


### c.
```{r}
summary(polr(response.fac ~ Male * Alternating, data=freq.table2))
```

```{r}
1-pchisq(789.0566-788.0099,1)
```

Since the value is greater than 0.05, this suggests that the interaction term is
not beneficial to the model in this instance.