---
title: "Assignment 4"
author: "Mason Rocco"
date: "2023-04-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(scatterplot3d)
```

# Problem 1

### a)
```{r, message=FALSE}
bone_length1 <- read_table("C:/Users/mrocc/Desktop/STAT 3302/Data/bone_length.txt")
bone_length <- bone_length1[,-1]
y_bar <- round(colMeans(bone_length),2)
print(y_bar)
S <- round(cov(bone_length),2)
print(S)
R <- round(cor(bone_length),2)
print(R)
```

### b)
```{r, message=FALSE}
det_S <- det(S)
print(det_S)
```

### c)
```{r, message=FALSE}
par(mfrow=c(1,1))
matplot(t(bone_length), type="l", xaxt="n", col="gray",
        xlab="Variable", ylab="Measurement")
axis(side=1, at=1:4, names(bone_length))
points(1:4, colMeans(bone_length), cex=1, pch=20, type="b", lwd=2)
```

# Problem 2

### a)
```{r, message=FALSE}
blood_glucose <- read_table("C:/Users/mrocc/Desktop/STAT 3302/Data/blood_glucose.txt")

means <- round(colMeans(blood_glucose),2)
y_bar2 <- means[1:3]
print(y_bar2)
x_bar2 <- means[4:6]
print(x_bar2)

S2 <- round(cov(blood_glucose),2)
Syy <- S2[1:3,1:3]
print(Syy)

Syx <- S2[1:3,4:6]
print(Syx)

Sxy <- S2[4:6,1:3]
print(Sxy)

Sxx <- S2[4:6,4:6]
print(Sxx)
```


### b)
```{r, message=FALSE}
R2 <- cor(blood_glucose)
print(R2)
pairs(blood_glucose)
```

As can be noted by both the correlation matrix and the pairwise scatterplots
of the data, there is not any correlation that is very strong between the 
variables. The strongest correlations are between the observations that took
place at the same occasion. For example, we see a slight correlation between 
y1 and x1, as well as between y2 and x2, and the same between y3 and x3.
